---
title: "Data"
output: html_document
---

```{r, include = FALSE}
library(survivoR)
library(tidyverse)
library(stringr)
library(dplyr)
library(survival)
library(survminer)
library(gtsummary)
library(gridExtra)
library(RColorBrewer)
library(usmap)
library(plotly)
library(knitr)
library(broom)
```

```{r, warning = FALSE, message = FALSE}
survivor_data_final = read.csv("data/survivor_data_final.csv")
```

## Data source

I used data sources from the `castaway_details` and `castaways` datasets from the [*survivoR*](https://github.com/doehm/survivoR.git) package. This package contains data from all seasons of Survivor. After cleaning, the final dataset contains unique information for each castaway for each season, including the following key variables:

**`version_season`**: version and season number\
**`full_name`**: contestant full name\
**`age_during_show`**: age, in years\
**`race`**: Race of castaway\
**`ethnicity`**: Ethnicity of castaway\
**`gender`**: 2 levels: Female, Male.\
**`personality_type_binary`**: 2 levels: Extrovert, Introvert.\
**`days_survived`**: Number of days survived in the show until elimination\
**`region`**: region in the U.S. where the contestant is from. I created this variable based on the state variable available in the dataset. 4 levels: West, Midwest, Northeast, South.\

## Summary

```{r, warning = FALSE, message = FALSE}
survivor_data_final %>% 
  select(gender, race, personality_type_binary, age_during_show, days_survived, region) %>% 
  tbl_summary(type = list(gender ~ "categorical", 
                          race ~ "categorical", 
                          personality_type_binary ~ "categorical", 
                          region ~ "categorical",
                          age_during_show ~ "continuous", 
                          days_survived ~ "continuous"), 
              statistic = list(all_continuous() ~ "{mean} ({sd})"), 
              digits = all_continuous() ~ 1,
              label = list(c(gender) ~ "Gender",
                           c(race) ~ "Race Identifier",
                           c(personality_type_binary) ~ "Personality Type",
                           c(region) ~ "Region",
                           c(age_during_show) ~ "Age During Show (Years)",
                           c(days_survived) ~ "Survival Time on Show (Days)")) %>% 
  bold_labels()
``` 

Note: N = 852 refers to the total count of records (i.e. contestant occurrences) in `survivor_data_final`; distinct persons may be listed in multiple records, across seasons and/or within seasons. 


## Mean Days Survived, by Contestant Characteristics{.tabset}

### Personality Type
```{r, warning = FALSE, message = FALSE}
## Personality type
# p-value
personality_p_value = t.test(days_survived ~ personality_type_binary, data = survivor_data_final) %>% 
  tidy() %>% 
  pull(p.value)

survivor_data_final %>%
  group_by(personality_type_binary) %>%
  summarize(
    n_personality_dist = n_distinct(full_name),
    n_personality_occ = n(),
    mean_days_survived = mean(days_survived, na.rm = TRUE)
  ) %>%
  na.omit() %>%
  mutate(`p-value` = ifelse(row_number() == 1, round(personality_p_value, 3), "")) %>%
  kable(digits = 1, col.names = c("Personality Type", "Distinct Persons", "Contestant Occurrences", "Mean Days Survived", "p-value"))
 
```

### Race
```{r}
## race
# p-value
race_aov = aov(days_survived ~ race, data = survivor_data_final)
race_p_value = summary(race_aov)[[1]][["Pr(>F)"]][1]

survivor_data_final %>%
  group_by(race) %>%
  summarize(
    n_race_dist = n_distinct(full_name),
    n_race_occ = n(),
    mean_days_survived = mean(days_survived, na.rm = TRUE)
  ) %>%
  na.omit() %>%
  mutate(`p-value` = ifelse(row_number() == 1, round(race_p_value, 3), "")) %>%
  kable(digits = 1, col.names = c("Race", "Distinct Persons", "Contestant Occurrences", "Mean Days Survived", "p-value"))
```

### Gender
```{r}
## Gender
# p-value
gender_p_value <- t.test(days_survived ~ gender, data = survivor_data_final) %>% 
  tidy() %>% 
  pull(p.value)

survivor_data_final %>%
  group_by(gender) %>%
  summarize(
    n_gender_dist = n_distinct(full_name),
    n_gender_occ = n(),
    mean_days_survived = mean(days_survived, na.rm = TRUE)
  ) %>%
  na.omit() %>%
  mutate(`p-value` = ifelse(row_number() == 1, round(gender_p_value, 3), "")) %>%
  kable(digits = 1, col.names = c("Gender", "Distinct Persons", "Contestant Occurrences", "Mean Days Survived", "p-value"))
```



## Gender Representation Across Seasons

```{r, warning = FALSE, message = FALSE}
fill_color = brewer.pal(9,"Purples")[5]
survivor_gender_over_time = survivor_data_final %>% 
  group_by(version_season, gender) %>%
  summarize(count = n_distinct(full_name)) %>%
  mutate(freq = count / sum(count)) %>%
  filter(gender == "Female") %>%
  separate(col = version_season, into = c('NA', 'season'), sep = 2) %>%
  dplyr::select(-"NA") %>%
  mutate(season = as.numeric(season))

Women_proportion_plot = 
  ggplot(data = survivor_gender_over_time, aes(x = season, y = freq, group = 1)) +
  geom_line(color = fill_color) +
  geom_point(color = fill_color) +
  geom_smooth(se = FALSE, color = "purple") + 
  ggtitle("Change in Proportion of Women Over Time") +
  xlab("Season") + ylab("Proportion of Women")

ggplotly(Women_proportion_plot)
```

Note: Distinct person counts by gender.

## Concentration of Contestants by Age and Gender

```{r, warning = FALSE, message = FALSE}
fill_color = brewer.pal(3, "Set2")

dis_age_plot = ggplot(survivor_data_final, aes(x = age_during_show, fill = gender)) +
  geom_histogram(aes(y = after_stat(count/sum(count))), bins = 15, color = "white", position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = fill_color) +
  ggtitle("Distribution of Ages Across Contestant Occurrences by Gender") +
  xlab("Age (During Show)") +
  ylab("Count") +
  theme_minimal()

ggplotly(dis_age_plot)

```



```{r, warning = FALSE, message = FALSE}
survivor_state = survivor_data_final %>%
  group_by(state) %>%
  summarize(n = n())

map_plot =
  plot_usmap(
  data = survivor_state, values = "n", lines = "blue"
) + 
  scale_fill_continuous(type = "viridis", name = "Contestant Count", label = scales::comma
  ) + 
  labs(title = "US States", subtitle = "Geographic Distribution of Contestants") +
  theme(legend.position = "right")
ggplotly(map_plot)
```